### 2021-07-01 paper.tex
edits on Sec.4

\def\Nequals#1{$\mathcal{N}{=}#1$}
\def\bQ{\mathbb{Q}}
\def\bZ{\mathbb{Z}}
\def\MF{\mathrm{MF}}
\def\TMF{\mathrm{TMF}}
\def\Tmf{\mathrm{Tmf}}
\def\tmf{\mathrm{tmf}}
\def\tr{\mathop{\mathrm{tr}}}

@@

\section{Fermion contribution to anomalies}
\label{sec:fermion}
So far, we first determined the 2-group structure in Sec.~\ref{sec:2-group} by studying the charges of line operators,
and then determined the anomalies in Sec.~\ref{sec:sl2z} by matching it to the action of $\SL(2,\bZ_2)$.
Going over  the entries on the column $\Spin$ of Table~\ref{table:refined},
we find that the anomaly is trivial when $(n_c,n_f)$ is (even,even) or (odd,even),
while it is $\alpha$ when $(n_c,n_f)$ is (odd,even) and $\alpha'$ when (odd,odd).
Since the 1-form symmetry background in the $\Spin$ theory is simply the Stiefel-Whitney class $w_2$ of the $\SO(2n_c)$ gauge bundle, 
these anomalies should simply come from the anomalies of fermions charged under $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$.
Here we use $\USp(2n_f)$ instead of $\SU(2n_f)$, because under the latter we also have perturbative anomalies, which might complicate the analysis.

For even $n_c$, the anomaly should be given by  \begin{equation}
\alpha= 2\pi i \cdot \frac12 w_2 \beta a_2
\end{equation}
where $w_2,a_2\in H^2(X;\bZ_2)$ controls the lift from $\SO(2n_c)$ to $\Spin(2n_c)$
and from $\USp(2n_f)/\bZ_2$ to $\USp(2n_f)$, respectively.
For odd $n_c$, the anomaly cochain should be given by 
 \begin{equation}
\alpha'=2\pi i \cdot \frac12 v_2 \beta_2 v_2
\end{equation} where $v_2$ is the class in $H^2(X;\bZ_2)$ which is the mod-2 reduction of
the class $x_2 \in H^2(X;\bZ_4)$ controlling the lift from $\SO(2n_c)/\bZ_2=\Spin(2n_c)/\bZ_4$ to $\Spin(2n_c)$.
We note that $\alpha'$ is exact, but as explained in the previous section, this still affects the gauging process.

@@

Before proceeding, we have two remarks.
First, this method was first used in \cite[Sec.~4]{Witten:1995gf} under the name of `a curious minus sign', and recognized as determining an anomaly in \cite[Sec.~2.4.3]{Cordova:2018acb}.
It was also used in \cite[Sec.~3.1 and Sec.~5.1.2]{Wang:2018qoy}.
Second, in this section we can only say that the effective interaction we find is compatible with the anomalies
as found in Sec.~\ref{sec:sl2z},
and will not be able to determine it completely.
This is mostly due to the fact that the computation of the spin bordism group $\Omega^\text{spin}_d([\SO(2n_c)\times \USp(2n_f)]/\bZ_2)$ which governs the anomaly is quite hard, 
because even the integral cohomology of the group in question is hard to obtain.
Only in a couple of cases we can say more, as we comment along the way.

\subsection{Effective interaction}
We break $\USp(2n_f)$ down to $\U(n_f)$ using a scalar field,
such that the fundamental representation of $\USp(2n_f)$ splits to the fundamental plus the anti-fundamental representation of $\U(n_f)$.
The monopole charge is given by the first Chern class $c_1$ of the low-energy $\U(n_f)$ flavor symmetry.

Take a standard 't Hooft-Polyakov monopole associated to $\U(1)\subset \USp(2)$
and embed it into  $\U(n_f)\subset \USp(2n_f)$.
The fermion zero modes form a vector representation of $\SO(2n_c)$,
whose quantization leads to the spinor representation.
This means that there is an effective interaction \begin{equation}
2\pi i \cdot \frac12 \int_X w_2 c_1.
\end{equation} %where $w_2$ is the Stiefel-Whitney class of the $\SO(2n_c)$ bundle
%and $c_1$ is the first Chern class of $\U(n_f)$.

\subsection{Anomalies}
We now take the $\bZ_2$ quotient, changing the symmetry group from 
$\SO(2n_c)\times \USp(2n_f)$ to $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$.
Note that $\pi_1(U(n_f)/\bZ_2)=\bZ\times \bZ_2$  or $\bZ$ depending on whether $n_f$ is even or odd.

Furthermore, when $n_f$ is even, $c_1(\U(n_f)) = c_1(\U(n_f)/\bZ_2)$ and $a_2(\U(n_f)/\bZ_2)=a_2(\USp(2n_f))$.
When $n_f$ is odd, $c_1(\U(n_f)/\bZ_2)=2 c_1(\U(n_f))$ when the latter is well-defined and $a_2(\U(n_f)/\bZ_2)$ is the mod-2 reduction of $c_1(\U(n_f)/\bZ_2)$.
We now compute the anomaly cochains in the four cases separately:
\paragraph{$(n_c,n_f)$=(even,even):}
$w_2(\SO(2n_c))$ and $c_1(\U(n_f))$ can be generalized without any problem,
and therefore \begin{equation}
\delta\left(\frac12 w_2(\SO(2n_c)) c_1(\U(n_f))\right) = 0.
\end{equation}

@@

\paragraph{$(n_c,n_f)$=(even,odd):}
Here we need to replace $c_1(\U(n_f))$ by $c_1(\U(n_f)/\bZ_2)/2$.
The effective interaction is then \begin{equation}
\frac14 w_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) 
\end{equation} and \begin{align}
\delta\left(\frac14 w_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right)
&= \frac12 \left(\frac12\delta w_2(\SO(2n_c)) c_1(\U(n_f)/\bZ_2) \right) \\
&= \frac12 \Big(\beta w_2(\SO(2n_c))\Big)c_1(\U(n_f)/\bZ_2) 
\end{align}
which is a pull-back of the anomaly cochain
\begin{equation}
 \frac12 \Big(\beta w_2(\SO(2n_c))\Big)a_2(\USp(2n_f)/\bZ_2). 
\end{equation}
When $n_c=2$ and $n_f=1$, we can confirm this is indeed the entire anomaly,
since we can compute $\mathrm{Hom}(\Omega^\text{spin}_5([\SO(4)\times \USp(2)]/\bZ_2),\U(1))$
and show that this is the only nontrivial element there.
For details, see Appendix~\ref{sec:so4su2}.

\paragraph{$(n_c,n_f)$=(odd,odd):}
Now we make the replacement on both sides and the effective interaction is \begin{equation}
\frac18 x_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) 
\end{equation}
and \begin{align}
\delta\left(\frac18 x_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right)
&= \frac12 \left(\frac14\delta v_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right) \\
&= \frac12 \Big(\beta_2 v_2(\SO(2n_c)/\bZ_2)\Big)c_1(\U(n_f)/\bZ_2) 
\end{align}
which is the pull-back of \begin{equation}
 \frac12 \Big(\beta_2 v_2(\SO(2n_c))\Big)a_2(\USp(2n_f)).
\end{equation}
We now recall that the symmetry we are considering now is $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$,
and therefore there is a single degree-2 obstruction cochain which equals both $v_2$ and $a_2$,
and therefore the anomaly cochain is \begin{equation}
 \frac12 v_2\beta_2 v_2.
\end{equation}
This is what we wanted to show.
Recall that this anomaly cochain is exact as we have repeatedly mentioned.
This is consistent with our computation of the bordism group in Appendix~\ref{sec:so(2odd)},
where we show that $\Omega^\text{spin}_5([\SO(2n_c)\times \USp(2n_f)]/\bZ_2=0$ when $n_c$ is odd.

>>>>

\def\Nequals#1{$\mathcal{N}{=}#1$}
\def\bR{\mathbb{R}}
\def\bQ{\mathbb{Q}}
\def\bZ{\mathbb{Z}}
\def\MF{\mathrm{MF}}
\def\TMF{\mathrm{TMF}}
\def\Tmf{\mathrm{Tmf}}
\def\tmf{\mathrm{tmf}}
\def\tr{\mathop{\mathrm{tr}}}

@@

\section{Fermion contribution to anomalies}
\label{sec:fermion}
So far, we first determined the 2-group structure in Sec.~\ref{sec:2-group} by studying the charges of line operators,
and then determined the anomalies in Sec.~\ref{sec:sl2z} by matching it to the action of $\SL(2,\bZ_2)$.
Going over  the entries on the column $\Spin$ of Table~\ref{table:refined},
we find that the anomaly is trivial when $(n_c,n_f)$ is (even,even) or (odd,even),
while it is $\alpha$ given in \eqref{anom} when $(n_c,n_f)$ is (odd,even) 
and $\alpha'$ given in \eqref{anom'} when (odd,odd).
Since the 1-form symmetry background in the $\Spin$ theory is simply the Stiefel-Whitney class $w_2$ of the $\SO(2n_c)$ gauge bundle, 
these anomalies should simply come from the anomalies of fermions charged under $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$.
Here we use $\USp(2n_f)$ instead of $\SU(2n_f)$, because under the latter we also have perturbative anomalies, which would complicate the analysis.

For even $n_c$, the anomaly should be given by  \begin{equation}
\alpha=% 2\pi i \cdot 
\frac12 w_2 \beta a_2
\end{equation}
where $w_2,a_2\in H^2(X;\bZ_2)$ controls the lift from $\SO(2n_c)$ to $\Spin(2n_c)$
and from $\USp(2n_f)/\bZ_2$ to $\USp(2n_f)$, respectively.
For odd $n_c$, the anomaly cochain should be given by 
 \begin{equation}
\alpha'=%2\pi i \cdot
 \frac12 v_2 \beta_2 v_2
\end{equation} where $v_2$ is the class in $H^2(X;\bZ_2)$ which is the mod-2 reduction of
the class $x_2 \in H^2(X;\bZ_4)$ controlling the lift from $\SO(2n_c)/\bZ_2=\Spin(2n_c)/\bZ_4$ to $\Spin(2n_c)$.
We note that $\alpha'$ is exact, but as explained in the previous section, this still affects the gauging process.

@@

Before proceeding, we have two remarks.
First, this method was first used in \cite[Sec.~4]{Witten:1995gf} to understand `a curious minus sign' 
appearing in the topologically-twisted Seiberg-Witten theory, 
which was more recently recognized as determining an anomaly in \cite[Sec.~2.4.3]{Cordova:2018acb}.
It was also used in \cite[Sec.~3.1 and Sec.~5.1.2]{Wang:2018qoy} to relate the `new' $\SU(2)$ anomaly 
with the effective interaction in the $\U(1)$ theory.
Second, in this section we can only say that the effective interaction we find is compatible with the anomalies
as found in Sec.~\ref{sec:sl2z},
and will not be able to determine it completely.
This is mostly due to the fact that the computation of the spin bordism group $\Omega^\text{spin}_d([\SO(2n_c)\times \USp(2n_f)]/\bZ_2)$ which governs the anomaly is quite hard, 
because even the integral cohomology of the group in question is hard to obtain.
Only in a couple of cases we can say more, as we comment along the way.

\subsection{Effective interaction}
We break $\USp(2n_f)$ down to $\U(n_f)$ using a scalar field,
such that the fundamental representation of $\USp(2n_f)$ splits into the fundamental plus the anti-fundamental representation of $\U(n_f)$.
The monopole charge is given by the first Chern class $c_1$ of the low-energy $\U(n_f)$ flavor symmetry.

Take a standard 't Hooft-Polyakov monopole associated to $\U(1)\subset \USp(2)$
and embed it into  $\U(n_f)\subset \USp(2n_f)$.
The fermion zero modes form a vector representation of $\SO(2n_c)$,
whose quantization leads to the spinor representation.

As first discussed in \cite{Thorngren:2014pza} and also used in \cite[Sec.~3.1]{Wang:2018qoy},
this means that there is an effective interaction \begin{equation}
%2\pi i \cdot
 \frac12 \int_X w_2 c_1.
 \label{bulk}
\end{equation} %where $w_2$ is the Stiefel-Whitney class of the $\SO(2n_c)$ bundle
%and $c_1$ is the first Chern class of $\U(n_f)$.
One way to understand it is as follows.

We started from a system which has $\SO(2n_c)$ symmetry,
but the spinor representation is only a projective representation of this symmetry.
This is an anomaly at the core of the monopole, which needs to flow in from the bulk.
Indeed, take the spacetime  to be $X=\bR_{\ge 0} \times \bR_t \times S^2$ around the monopole.
Reducing the bulk term \eqref{bulk} on $S^2$ with $\int_{S^2} c_1=1$, 
we have the  effective interaction $\int_Y w_2$
on the half-space $Y=\bR_{\ge 0}\times \bR_t$,
and the monopole lives on the boundary. 
Therefore, the degree of freedom on the boundary is in the projective representation characterized by 
$w_2\in H^2(\SO(2n_c),\bZ_2)$.

\subsection{Anomalies}
We now take the $\bZ_2$ quotient, changing the symmetry group from 
$\SO(2n_c)\times \USp(2n_f)$ to $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$.
Note that $\pi_1(\U(n_f)/\bZ_2)=\bZ\times \bZ_2$  or $\bZ$ depending on whether $n_f$ is even or odd.
We denote by $a_2$ the obstruction class to lift a $\U(n_f)/\bZ_2$ bundle to a $\U(n_f)$ bundle.
This implies the following:
\begin{itemize}
\item When $n_f$ is even, $c_1(\U(n_f)) = c_1(\U(n_f)/\bZ_2)$ and $a_2(\U(n_f)/\bZ_2)=a_2(\USp(2n_f))$.
\item When $n_f$ is odd, $c_1(\U(n_f)/\bZ_2)=2 c_1(\U(n_f))$ when the latter is well-defined.
More generally, $a_2(\U(n_f)/\bZ_2)$ is the mod-2 reduction of $c_1(\U(n_f)/\bZ_2)$.
\end{itemize}
We now compute the anomaly cochains in the four cases separately:
\paragraph{$(n_c,n_f)$=(even,even):}
$w_2(\SO(2n_c))$ and $c_1(\U(n_f))$ can be generalized to closed cochains of $\SO(2n_c)/\bZ_2$ 
and of $\U(n_f)/\bZ_2$
without any problem,
and therefore \begin{equation}
\delta\left(\frac12 w_2(\SO(2n_c)) c_1(\U(n_f))\right) = 0.
\end{equation}

@@

\paragraph{$(n_c,n_f)$=(even,odd):}
Here we need to replace $c_1(\U(n_f))$ by $c_1(\U(n_f)/\bZ_2)/2$.
The effective interaction is then \begin{equation}
\frac14 w_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) 
\end{equation} and \begin{align}
\delta\left(\frac14 w_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right)
&= \frac12 \left(\frac12\delta w_2(\SO(2n_c)) c_1(\U(n_f)/\bZ_2) \right) \\
&= \frac12 \Big(\beta w_2(\SO(2n_c))\Big)c_1(\U(n_f)/\bZ_2) 
\end{align}
which is a pull-back of the anomaly cochain
\begin{equation}
 \frac12 \Big(\beta w_2(\SO(2n_c))\Big)a_2(\USp(2n_f)/\bZ_2). 
\end{equation}
When $n_c=2$ and $n_f=1$, we can confirm that this is indeed the entire anomaly,
since we can compute $\mathrm{Hom}(\Omega^\text{spin}_5([\SO(4)\times \USp(2)]/\bZ_2),\U(1))$
and show that this is the only nontrivial element there.
For details, see Appendix~\ref{sec:so4su2}.

\paragraph{$(n_c,n_f)$=(odd,odd):}
Now we make the replacement on both sides and therefore the effective interaction is \begin{equation}
\frac18 x_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) 
\end{equation}
and \begin{align}
\delta\left(\frac18 x_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right)
&= \frac12 \left(\frac14\delta v_2(\SO(2n_c)/\bZ_2) c_1(\U(n_f)/\bZ_2) \right) \\
&= \frac12 \Big(\beta_2 v_2(\SO(2n_c)/\bZ_2)\Big)c_1(\U(n_f)/\bZ_2) 
\end{align}
which is the pull-back of \begin{equation}
 \frac12 \Big(\beta_2 v_2(\SO(2n_c))\Big)a_2(\USp(2n_f)).
\end{equation}
We now recall that the symmetry we are considering now is $[\SO(2n_c)\times \USp(2n_f)]/\bZ_2$,
and therefore there is a single degree-2 obstruction cochain which equals both $v_2$ and $a_2$,
and therefore the anomaly cochain is \begin{equation}
 \frac12 v_2\beta_2 v_2.
\end{equation}
This is what we wanted to show.
Recall that this anomaly cochain is exact as we have repeatedly mentioned.
This is consistent with our computation of the bordism group in Appendix~\ref{sec:so(2odd)},
where we show that $\Omega^\text{spin}_5([\SO(2n_c)\times \USp(2n_f)]/\bZ_2=0$ when $n_c$ is odd.
